---
title: "Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data

This is a project to explore activity monitoring data. We will import the data to a datafram

```{r import, echo=TRUE}
df=read.csv('D:/temp/repdata_data_activity/activity.csv')
df$date=as.Date(df$date,"%Y-%M-%d")
head(df)

```
##Calculate mean 

We will now calculate the mean number of steps taken per daytot

```{r mean, echo=TRUE}
totalsteps <- tapply(df$steps, df$date, sum)
mean(totalsteps, na.rm = TRUE )
median(totalsteps, na.rm = TRUE )

hist (totalsteps, breaks = 10)

```

##Generate average steps time series
We will generate a plot showing average steps per 5 min time period.

```{r time series plot, echo=TRUE}
averagesteps <- aggregate(df$steps, by=list(df$interval), FUN = mean, na.rm = TRUE)

plot(x=averagesteps$Group.1, y=averagesteps$x, type = 'l', xlab = 'Time', ylab = 'Average Steps', main = 'Average Steps Time Series')

averagesteps[which.max(averagesteps$x),1]

```

##Impute missing numbers
We will calculate and impute missing numbers

``` {r missing numbers, echo= TRUE}

##number of incompletecases
sum(!complete.cases(df))
dfcopy <- df
library(dplyr)
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
dfcopy<-dfcopy %>%
  group_by(interval) %>%
  mutate(
    steps=impute.mean(steps)
  )

totalsteps <- tapply(dfcopy$steps, dfcopy$date, sum)
mean(totalsteps)
median(totalsteps)

hist (totalsteps, breaks = 10)

```

#Weekends vs weekdays
We will see if there is difference between activity pattern on weekdays and weekends

``` {r weekends, echo=TRUE}

dfcopy$weekend <- 'Weekday'

for (i in 1:nrow(dfcopy)){
  if (weekdays(dfcopy$date[i])=="Saturday" | weekdays(dfcopy$date[i])=="Sunday" ) {
    dfcopy$weekend[i]<-'Weekend'
  }
}
averagesteps <- aggregate(x=list(steps=dfcopy$steps), by=list(interval=dfcopy$interval, weekend=dfcopy$weekend), FUN = mean, na.rm = TRUE)
library(ggplot2)
qplot(x=interval, y=steps, data=averagesteps, geom="line", facets = weekend~.)

```
